Effects of Different Pain Relievers and Dosages on Cognitive Retention

### 2025-03-11 paper/paper.qmd
Assumptions

>>>>

**Repeated Measures ANOVA**
(To be changed)
When performing a repeated measures ANOVA, it is essential to verify that the underlying assumptions of the test are met. These assumptions include:
\begin{itemize}
    \item \textbf{Normality}: The dependent variable should be normally distributed within each group.
    \item \textbf{Sphericity}: The variances of the differences between all combinations of related groups should be equal.
    \item \textbf{Homogeneity of Variance}: The variances of the dependent variable should be equal across groups.
\end{itemize}

This document outlines the methods and tests used to verify these assumptions.

\section*{Normality}
The assumption of normality was tested using the Shapiro-Wilk test. The Shapiro-Wilk test evaluates the null hypothesis that the data is normally distributed. The test statistic \( W \) and the corresponding \( p \)-value were calculated for each combination of the factors (Drug, DosageLevel, and Time).

\subsection*{Hypotheses}
\begin{itemize}
    \item \( H_0 \): The data is normally distributed.
    \item \( H_1 \): The data is not normally distributed.
\end{itemize}

\subsection*{Interpretation}
If the \( p \)-value is greater than 0.05, we fail to reject the null hypothesis and assume normality. Otherwise, the normality assumption is violated.

\subsection*{Visual Inspection}
Q-Q plots were generated to visually inspect the normality assumption. If the points in the Q-Q plot lie approximately on the reference line, the data can be assumed to be normally distributed.

\section*{Sphericity}
The assumption of sphericity was tested using Mauchly's test. Sphericity refers to the condition where the variances of the differences between all combinations of related groups are equal.

\subsection*{Hypotheses}
\begin{itemize}
    \item \( H_0 \): The variances of the differences are equal (sphericity is met).
    \item \( H_1 \): The variances of the differences are not equal (sphericity is violated).
\end{itemize}

\subsection*{Interpretation}
If the \( p \)-value from Mauchly's test is greater than 0.05, we fail to reject the null hypothesis and assume sphericity. If the assumption is violated, corrections such as Greenhouse-Geisser or Huynh-Feldt should be applied.

\section*{Homogeneity of Variance}
The assumption of homogeneity of variance was tested using Levene's test. This test evaluates whether the variances of the dependent variable are equal across all groups formed by the combinations of the factors (Drug, DosageLevel, and Time).

\subsection*{Hypotheses}
\begin{itemize}
    \item \( H_0 \): The variances are equal across groups.
    \item \( H_1 \): The variances are not equal across groups.
\end{itemize}

\subsection*{Interpretation}
If the \( p \)-value from Levene's test is greater than 0.05, we fail to reject the null hypothesis and assume homogeneity of variance. If the assumption is violated, transformations or alternative statistical methods may be required.

\subsection*{Visual Inspection}
Boxplots were generated to visually inspect the homogeneity of variance. If the boxplots show similar spreads across groups, the assumption is likely met.

\section*{Conclusion}
The results of the assumption tests (Shapiro-Wilk, Mauchly's, and Levene's) were used to determine whether the assumptions of normality, sphericity, and homogeneity of variance were met. If any assumptions were violated, appropriate corrections or alternative methods were considered before proceeding with the repeated measures ANOVA.
```{r}
#| label: tbl-model1
#| tbl-cap: "Summary of Age Groups"
#| echo: false
#| warning: false
#| message: false

# Load necessary libraries
library(tidyverse)
library(ggplot2)
library(car)
library(ez)

# Assuming the dataset is loaded into a data frame called `data`
# Convert necessary columns to factors
clean_data$Drug <- as.factor(clean_data$Drug)
clean_data$DosageLevel <- as.factor(clean_data$DosageLevel)
clean_data$Time <- as.factor(clean_data$Time)
clean_data$AgeGroup <- as.factor(clean_data$AgeGroup)

# Check for Normality
# We will use Shapiro-Wilk test for normality
shapiro_test <- clean_data %>%
  group_by(Drug, DosageLevel, Time) %>%
  summarise(statistic = shapiro.test(GameScore)$statistic,
            p.value = shapiro.test(GameScore)$p.value)

print(shapiro_test)

# Visual check for normality using Q-Q plots
ggplot(clean_data, aes(sample = GameScore)) +
  geom_qq() +
  geom_qq_line() +
  facet_grid(Drug ~ DosageLevel + Time) +
  theme_minimal() +
  ggtitle("Q-Q Plots for Normality Check")

# Check for Sphericity using Mauchly's Test
# We will use the `ezANOVA` function to check for sphericity
ez_model <- ezANOVA(
  data = clean_data,
  dv = .(GameScore),
  wid = .(Name),
  within = .(Time),
  between = .(Drug, DosageLevel),
  type = 3,
  detailed = TRUE
)

print(ez_model)

# Check for Homogeneity of Variance using Levene's Test
levene_test <- leveneTest(GameScore ~ Drug * DosageLevel * Time, data = clean_data)
print(levene_test)

# Visual check for homogeneity of variance using boxplots
ggplot(clean_data, aes(x = interaction(Drug, DosageLevel, Time), y = GameScore)) +
  geom_boxplot() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Boxplots for Homogeneity of Variance Check")
